Optimal Particle Paths Around Points in $\R^2$

### 2013-09-28 particle_definitions.tex
Add more proofs.

@@

\begin{definition}
  For a given particle, $p$, a fastest path, $\hat{\gamma}(t)$, between two points, $\bvec{X_1}$ and $\bvec{X_2}$, is a valid path such that $T_f(\hat{\gamma}) \leq T_f(\gamma)$ for all valid paths, $\gamma(t)$, between $\bvec{X_1}$ and $\bvec{X_2}$.
\end{definition}

\begin{definition}
  The centripetal acceleration, $\bvec{a_c}$, of a particle, $p$, is the component of the acceleration of $p$ perpendicular to its direction of motion, $\bhat{v}$. The sign of $a_c$ is difined as the sign of the projection of $\bhat{a_c}$ onto $\bhat{r}$.
\end{definition}

\begin{definition}
  The tangential acceleration, $\bvec{a_t}$, of a particle, $p$, is the component of the acceleration of $p$ in its direction of motion, $\bhat{v}$.
\end{definition}

@@

The motion of the 2-dimensional particles in this paper will typically be described in the polar coordinate system, shown in figure \ref{fig:polar-param}.

@@

@@

If we look at a point, $p$, subject to only a centripetal acceleration, $a_c$, the change in $\bvec{v}$ over an infinitesimal time, $dt$, is shown in the figure \ref{fig:theta-deriv} (the two vectors, $\bvec{v}$ and $\bvec{v}+\bvec{dv}$, are superimposed). From the definition of $a_c$

@@

So...
\[
\frac{d\beta}{dt} = \frac{a_c}{s}
\]
\end{proof}

@@

Returning again to figure \ref{fig:polar-param}, the following equations can be derived

\begin{align}
  \frac{dr(t)}{dt}& = -s(t) \, \cos\paren{\theta(t)}\\
  \frac{d\phi(t)}{dt}& = s(t) \, \sin\paren{\theta(t)}\\
  \frac{d\theta(t)}{dt}& = \frac{d\phi(t)}{dt} - \frac{a_c(t)}{s(t)}\label{eq:theta-deriv}\\
  \frac{d}{dt} \frac{dr(t)}{dt}& = -\frac{ds(t)}{dt} \cos\paren{\theta(t)} + s(t) \sin\paren{\theta(t)} \frac{d\theta(t)}{dt}
\end{align}

@@

  (1) The functions $\phi(t)$ and $\theta(t)$ are continuous.

  (2) For two times $t_1$ and $t_2$, s.t. $t_1 < t_2$, $\theta(t_1) > a$ and $\theta(t_2) < a$, then $\theta(t_c) = a$ for some $t_c \in [t_1, t_2]$.
\end{lemma}

\begin{proof}
First off, it should be noted that a solution to \ref{eq:theta-deriv} exists because the rhs is Lipschitz continuous. The proof of (1) follows directly from the fact that the derivative of $\theta(t)$ exist and is bounded. (2) is just a restatement of the intermediate value theorem.
\end{proof}

@@

\begin{proof}
The second case is the easiest to check,

>>>>

In this section, we will provide some basic definitions about particles and paths. These will lay the groundwork for thinking about optimal paths. We will begin by defining the description of a particle and then define various types of paths along which a particle can travel.

@@

\begin{definition}
  For a given particle, $p$, a fastest path, $\hat{\gamma}(t)$, between two points, $\bvec{X_1}$ and $\bvec{X_2}$, is a valid path such that $T_{f,\hat{\gamma}} \leq T_{f,\gamma}$ for all valid paths, $\gamma(t)$, between $\bvec{X_1}$ and $\bvec{X_2}$.
\end{definition}

\begin{definition}
  A particle's velocity is defined as

  \[\bvec{v} \coloneqq \frac{d\bvec{X}}{dt}\]

  A particle's speed is $v$, and its direction of motion is $\bhat{v}$

\end{definition}

\begin{definition}
  A particle's acceleration is defined as

  \[\bvec{a} \coloneqq \frac{d\bvec{v}}{dt}\]
\end{definition}

\begin{definition}
  The centripetal acceleration, $\bvec{a_c}$, of a particle, $p$, is the component of the acceleration of $p$ perpendicular to its direction of motion.

  In rectangular coordinates, the sign of $a_c$ is the projection of $\bhat{a_c}$ onto $\bhat{x}$.

  In polar coordinates, the sign of $a_c$ is difined as the sign of the projection of $\bhat{a_c}$ onto $\bhat{r}$.
\end{definition}

\begin{definition}
  The tangential acceleration, $\bvec{a_t}$, of a particle, $p$, is the component of the acceleration of $p$ in its direction of motion, $\bhat{v}$. $a_t = \frac{ds}{dt}$.
\end{definition}

@@

The motion of the 2-dimensional particles in this paper will be described in the polar coordinate system, shown in Figure \ref{fig:polar-param}.

@@

If we look at a point, $p$, subject to only centripetal acceleration, $a_c$, the change in $\bvec{v}$ over an infinitesimal time, $dt$, is shown in Figure \ref{fig:theta-deriv} (the two vectors, $\bvec{v}$ and $\bvec{v}+\bvec{dv}$, are superimposed).

@@

From the definition of $a_c$

@@

So...
\[
\frac{d\beta}{dt} = \frac{a_c}{v}
\]
\end{proof}

@@

Returning again to Figure \ref{fig:polar-param}, the following equations can be derived

\begin{align}
  \frac{dr(t)}{dt}& = -s(t) \, \cos\paren{\theta(t)}\label{eq:r-deriv}\\
  \frac{d\phi(t)}{dt}& = s(t) \, \sin\paren{\theta(t)}\label{eq:phi-deriv}\\
  \frac{d\theta(t)}{dt}& = \frac{d\phi(t)}{dt} - \frac{a_c(t)}{s(t)}\\
  &= s(t) \, \sin\paren{\theta(t)} - \frac{a_c(t)}{s(t)}\label{eq:theta-deriv}\\
  \intertext{Applying the chain rule to (\ref{eq:r-deriv})}
  \frac{d}{dt} \frac{dr(t)}{dt}& = -\frac{ds(t)}{dt} \cos\paren{\theta(t)} + s(t) \sin\paren{\theta(t)} \frac{d\theta(t)}{dt}\\
  &\begin{aligned}
    = -\frac{ds(t)}{dt} \cos\paren{\theta(t)} &+ s(t)^2 \sin^2\paren{\theta(t)}\\
    &\qquad - s(t) \sin\paren{\theta(t)} \frac{a_c(t)}{s(t)}\label{eq:r-deriv-2}
  \end{aligned}
\end{align}

@@

  1. The functions $\phi(t)$ and $\theta(t)$ are continuous.

  2. For two times $t_1$ and $t_2$, s.t. $t_1 < t_2$, $\theta(t_1) > a$ and $\theta(t_2) < a$, then $\theta(t_c) = a$ for some $t_c \in [t_1, t_2]$.
\end{lemma}

\begin{proof}
First off, it should be noted that a solution to \ref{eq:theta-deriv} exists because the right hand side is Lipschitz continuous. The proof of 1.\ follows directly from the fact that the derivative of $\theta(t)$ exist and is bounded. 2.\ is just a restatement of the intermediate value theorem.
\end{proof}

@@

For the first case

The second case is the easiest to check. If we plug $\theta(t_0) = 0$ into \ref{eq:theta-deriv}, we get

\begin{align*}
\frac{d\theta(t)}{dt} = 0 \qquad \text{for} \;\; t \ge t_0\\
\intertext{So...}
\theta(t) = 0 \qquad \text{for} \;\; t \ge t_0
\end{align*}

### 2013-09-28 path_theorems.tex
Add more proofs.

\begin{lemma}
If $a_t=0$ and $\norm{a_c} \le a_{c,max}$, then the minimum radius of curvature of a point trajectory is given by
\begin{equation}
R_{min} = \frac{v^2}{a_{c,max}}
\end{equation}
\end{lemma}

>>>>

\begin{lemma}
If $a_t=0$ and $\norm{a_c} \le a_{c,max}$, then the minimum radius of curvature of a point's trajectory is given by
\begin{equation}
R_{min} = \frac{v^2}{a_{c,max}}
\end{equation}
\end{lemma}

\begin{proof}
This comes directly from the definition of radius of curvature as

\[
R = \frac{v^2}{a_c}
\]
\end{proof}
